---
title: "Loewen et al. Exploratory Analysis"
author: "Spencer Braun"
date: "11/5/2020"
output: html_document
---



```{r, message=F}
library(DOS2)
library(optmatch)
library(RItools)
library(rcbalance)

library(readstata13)
library(tidyverse)
library(haven)
library(sandwich)
library(lmtest)
library(multiwayvcov)

source("utility.R")
```


```{r}
ces.raw <- read_dta("original_data/CES_DATA.dta")
main.raw <- read_dta("original_data/Main\ Data.dta")
reoffering.raw <- read.dta13("original_data/Reoffering\ Data.dta")
```

```{r}
View(ces.raw)
View(main.raw)
View(reoffering.raw)

```

### Exploratory

```{r}
main.raw %>% names()
```

```{r}
ces.raw %>% head()
```



```{r}
reoffering.raw %>%
  filter(placeonlist != place2006)

reoffering.raw %>%
  group_by(mp) %>%
  summarise(count=n())

reoffering.raw %>%
  filter(reoffered == 1) %>%
  filter(placeonlist != place2006)
```


### Main Analysis 
```{r}
main <- main.raw %>%
  filter(excluded != 1) %>%
  mutate(senate = ifelse(is.na(senate), 0, 1)) %>%
  mutate(optout = case_when(
    (place_on_paper <= 87) & (election == 2006) & (check == 0) ~ 1, #guessing place on paper = place from do file 
    (place_on_paper < 117) & (election == 2008) & (check == 0) ~ 1,
    T ~ 0
  )) %>%
  mutate(p2p = case_when(
     (place_on_paper <= 87) & (election == 2006) ~ 1,
     (place_on_paper < 117) & (election == 2008) ~ 1,
     T ~ 0
  )) %>%
  mutate(govp2p = gov * p2p) %>%
  mutate(transfersshare = transfers / limit) %>%
  mutate(transfersshare = ifelse(is.na(transfersshare) & !is.na(spending), 0, transfersshare)) %>%
  mutate(donationsshare = amount / limit) %>%
  mutate(donationsshare = ifelse(is.na(donationsshare) & !is.na(spending), 0, transfersshare)) %>%
  mutate(share = ifelse(!is.na(transfersshare + donationsshare) & (transfersshare + donationsshare > 1), 1, transfersshare + donationsshare))
  

write.csv(main, "processed_data/main.csv")
# main.raw %>% names() %>% sort()
# main.raw %>% select(order_of_precedence)
# main.raw %>% select(place_on_paper) 
```
```{r}
View(main)
```


Table 1
```{r}
table1.fit <- lm(current_vote ~ govp2p + gov + p2p + previous_vote + y2006, data= main)
summary(table1.fit)

# Note the paper uses clustered standard errors on member ID.
# Output below matches SE's and p values of table 1
vcov_firm <- cluster.vcov(table1.fit,main$id)
coeftest(table1.fit, vcov_firm)
```

```{r}
# correct filters since N matches
main %>% filter(gov == 0, y2006 == 0) %>% count()
main %>% filter(gov == 0, y2006 == 1) %>% count()
main %>% filter(gov == 1, y2006 == 0) %>% count()
main %>% filter(gov == 1, y2006 == 1) %>% count()
```
```{r}
mean((main %>% filter(gov == 1, y2006 == 1))$current_vote)
mean((main %>% filter(gov == 1, y2006 == 0))$current_vote)

```
 
The opportunity to introduce legislation increases the vote share of government candidates by 5.26 percentage points (p=0.01, two-tailed). It has no effectfor opposition members (p=0.57, two-tailed).

Matches our stata output, see Stata-MainResult.png. 
However p-values remain slightly different despite equal counts, hard to tell what else could be the reason besides different software defaults? Seems related to different DoF, probably not worth digging into 
```{r}
# t.test(current_vote ~ p2p, data = main %>% filter(gov == 1, y2006 == 1))
# t.test(current_vote ~ p2p, data = main %>% filter(gov == 1, y2006 == 0))

governing <- t.test(current_vote ~ p2p, data = main %>% filter(gov == 1))
governing

t.test(current_vote ~ p2p, data = main %>% filter(gov == 0))
main %>%
  group_by(p2p, gov) %>%
  summarise(count=n(), .groups='drop') %>%
  arrange(gov, p2p)
```

```{r}
governing$estimate[2] - governing$estimate[1]
governing$p.value
```



```{r}
main %>%
  select(p2p, current_vote, gov) %>%
  group_by(p2p, gov) %>%
  summarise(mean_vote = mean(current_vote)) %>%
  ggplot() + geom_bar(aes(x=factor(p2p), y=mean_vote, fill=factor(p2p)), stat='identity') +
  facet_wrap(.~factor(gov)) +
  scale_y_continuous(limits=c(0,100), breaks = seq(0,100,by=20)) + 
  theme_bw()
```


Mediation tests - see Stata-Mediation.png
```{r}
main %>% select(media_mentions)

t.test(media_mentions ~ p2p, data=main %>% filter(gov == 1))
```

```{r}
t.test(quality ~ p2p, data=main %>% filter(gov == 1))
```


This one differs from Stata results - look into how we are calculating.
```{r}
t.test(share ~ p2p, data=main %>% filter(gov == 1))
```

****SUPPORTING INFORMATION
```{r}

get_levels <- function(bys) main %>% select(!!!bys) %>% distinct() %>% pull(.)
sapply(get_levels("y2006"), function(level) t.test(p2p~ gov, data=main %>% filter(y2006 == level)))
```

```{r}
sapply(get_levels("gov"), function(level) t.test(p2p~ y2006, data=main %>% filter(gov == level)))
```

Randomization checks

```{r}
main %>% select(years_served)
t.test(years_served ~ p2p, data=main)
```

```{r}
main %>% select(exminister)
t.test(exminister ~ p2p, data=main)
```



### Reoffering Analysis

```{r}
reoffering <- reoffering.raw %>%
  mutate(p2p = case_when(
       (placeonlist <= 87) & (election == 2006) ~ 1,
       (placeonlist < 117) & (election == 2008) ~ 1,
       T ~ 0
    )) %>%
  mutate(place = case_when(
    election == 2006 ~ placeonlist / 237,
    election == 2008 ~ placeonlist / 255
  ))
```


```{r}
table.si4 <- glm(reoffered ~ p2p, data=reoffering, family='binomial')
table.si4.2006 <- glm(reoffered ~ p2p, data=reoffering %>% filter(election == 2006), family='binomial')
table.si4.2008<- glm(reoffered ~ p2p, data=reoffering %>% filter(election == 2008), family='binomial')


summary(table.si4)
summary(table.si4.2006)
summary(table.si4.2008)
```


```{r}
table.si5 <- glm(reoffered ~ place, data=reoffering, family='binomial')
table.si5.2006 <- glm(reoffered ~ place, data=reoffering %>% filter(election == 2006), family='binomial')
table.si5.2008 <- glm(reoffered ~ place, data=reoffering %>% filter(election == 2008), family='binomial')


summary(table.si5)
summary(table.si5.2006)
summary(table.si5.2008)
```



### CES Analysis 

```{r}
ces <- ces.raw %>%
  mutate(id= row_number()) %>%
  mutate(panel = ifelse(!is.na(ces06_pes_a1) & !is.na(survey_type040608), 1, 0)) %>%
  mutate(election2006 = ifelse(is.na(ces06_pes_a1), 0, 1)) %>%
  mutate(election2008 = ifelse(is.na(ces08_pes_a1), 0, 1)) %>%
  mutate(local08 = case_when(
    ces08_pes_b8d == 5 ~ 0,
    ces08_pes_b8d == 8 ~ 0,
    ces08_pes_b8d == 1 ~ 1
  )) %>%
  mutate(local06 = case_when(
    ces06_pes_i1 == 5 ~ 0,
    ces06_pes_i1 == 8 ~ 0,
    ces06_pes_i1 == 1 ~ 1
  )) %>%
  mutate(locallike08 = case_when(
    ces08_pes_b8e == 2 ~ "CON",
    ces08_pes_b8e == 1 ~ "LIB",
    ces08_pes_b8e == 3 ~ "NDP",
    ces08_pes_b8e == 4 ~ "BQ"
  )) %>%
  mutate(locallike06 = case_when(
    ces06_pes_i2 == 2 ~ "CON",
    ces06_pes_i2 == 1 ~ "LIB",
    ces06_pes_i2 == 3 ~ "NDP",
    ces06_pes_i2 == 4 ~ "BQ"
  )) %>%
  mutate(incumbentlike06 = case_when(
    (lib2006 == 1) & (locallike06 != "LIB")~ 0,
    (lib2006 == 1) & (locallike06 == "LIB")~ 1
  )) %>%
  mutate(incumbentlike08 = case_when(
    (con2008 == 1) & (locallike08 != "CON")~ 0,
    (con2008 == 1) & (locallike08 == "CON")~ 1
  )) %>%
  mutate(gov = case_when(
    (lib2006 == 0) & (election2006 == 1) ~ 0,
    (con2008 == 0) & (election2008 == 1) ~ 0,
    (lib2006 == 1) & (election2006 == 1) ~ 0,
    (con2008 == 1) & (election2008 == 1) ~ 0
  ))


```


```{r}
ctab1 <- with(
  ces %>% filter(lib2006 == 1) %>% filter(election2006 ==1), 
  table(incumbentlike06, p2p2006, useNA = "ifany")
  )

chisq.test(ctab1)

ctab2 <- with(
  ces %>% filter(con2008 == 1) %>% filter(election2008 ==1), 
  table(incumbentlike08, p2p2008, useNA = "ifany")
  )

chisq.test(ctab2)

```

